Improving content discovery through combining linked data and data mining techniques

### 2015-12-05 report/report.tex
Written most of data pipeline definition.

Figure~\ref{fig:object-deref}

@@

        IRI -> RDFa [label="dereference"];
        RDFa -> RDF [label="parse"];
        RDF -> RDF1 [label="derefence/parse"];
        RDF -> RDF2 [label="dereference/parse"];
        RDF -> RDFp [label="\cup"];
        RDF1 -> RDFp [label="\cup"];
        RDF2 -> RDFp [label="\cup"];
        RDFp -> Features [label="content\_?p\_?v"];
      }
    \end{dot2tex}
  \end{center}
  \caption{Semantic web content miner with addtional dereferencing of linked entities}
\end{sidewaysfigure}

\begin{figure}[h]
  \begin{center}
    \begin{dot2tex}[dot,options=-t math,autosize,pgf]
      digraph g {
        rankdir=LR;

@@

        RDF -> Features [label="generate"];
      }
    \end{dot2tex}
  \end{center}
  \caption{Maximal web content miner}
\end{figure}

>>>>

Figure~\ref{fig:object-deref} shows the data pipeline with this additional
enrichment stage. Note that now we have potential for some stages being
executed in parallel.

@@

        IRI -> RDFa [label="dereference"];
        RDFa -> RDF [label="parse"];
        RDF -> RDF1 [label="derefence/parse"];
        RDF -> RDF2 [label="dereference/parse"];
        RDF -> RDFp [label="\cup"];
        RDF1 -> RDFp [label="\cup"];
        RDF2 -> RDFp [label="\cup"];
        RDFp -> Features [label="content\_?p\_?v"];
      }
    \end{dot2tex}
  \end{center}
  \caption{\label{fig:object.-deref}Semantic web content miner with addtional dereferencing of linked entities}
\end{sidewaysfigure}

So far in this section, we have inductively built up a data pipeline from
a ``null'' base working with only identity graphs to a simple pipeline
capable of \emph{extracting} an RDF graph, \emph{enriching} it and then
\emph{generating} features from it.

What needs to be proven through experimentation is \emph{which} of these
provides the information required for effective data mining. As part of this
experimentation, we can now look at further techniques and approaches to try.

\subsection{Improving Extraction}

In order to gain a larger set of RDF data at the start of the pipeline, we
can derive further ways to get information about a content item given only
its IRI as input.

Rizzo and Troncy\cite{rizzo2012nerd} defined a framework called NERD capable
of combining multiple entity extraction systems to provide a unified way of
identifying -- and disambiguating -- named entities within a given body of text.
With such a system, we can create a second, parallel RDF extraction strategy
that creates a graph of triples in the form:

$$
(IRI, rdf\!\!:\!\!about, Entity)
$$

\noindent where $Entity$ is an IRI representing a concept or entity believed
to be found in the content's textual content. A data pipeline complementing
the RDFa-based extraction is depicted in Figure~\ref{fig:entity-extraction}.
Note the ability to apply a simple set union to the result of each extraction
as with the enrichment.

\begin{sidewaysfigure}[h]
  \begin{center}
    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.8]
      digraph g {
        rankdir=LR;

        node [shape=circle,margin="0,0"]
        RDF1 [label="RDF_1"];
        RDF2 [label="RDF_2"];

        IRI -> RDFa [label="derefernce"];
        IRI -> Text [label="GET"];
        RDFa -> RDF1 [label="parse"];
        Text -> RDF2 [label="entity extraction"];
        RDF1 -> RDF [label="\cup"];
        RDF2 -> RDF [label="\cup"];
        RDF -> Features [label="content\_?p\_?v"];
      }
    \end{dot2tex}
  \end{center}
  \caption{Named entity extraction in addition to semantic web extraction\label{fig:entity-extraction}}
\end{sidewaysfigure}

Another strategy can be to infer a relationship between two content items where
one contents an HTML link to another. It is not always possible to derive
precise semantics of such a link (unless the publisher has kindly provided
a \texttt{rel} attribute), but a weak relationship such as:

$$
(IRI_1, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, IRI_2)
$$

\noindent might prove -- through experiementation -- to be useful enough
for data mining insights.

The fourth and final extraction strategy explored is acquiring metadata from
bespoke Content Management Systems and other internal APIs. This is generally
the only option used in enterprises settings, as discussed in
Section~\ref{sec:linked-enterprise-data}. This is likely to be the richest
source of information where an enterprise has typically preferred bespoke
integrations against non-hypermedia interfaces, so experimentation should
help quantity or qualify the value such a direct integration adds (perhaps to
consider it in combination with the cost of bepsoke, repeated integration
projects).

The assertion explored here is that such bespoke integrations can
\emph{complement} cheaper work such as extracting RDFa with pre-built tools
(and thus be developed one-by-one after the initial release of an application
such as this data pipeline). Note that repeated custom integration projects
means that each data source requires a different application be developed (as
opposed to the reuse of a single RDFa parser or HTML link scraper). This
also means we are not necessarily comparing like-for-like if we introduce only
one at a time. It also makes it
difficult to evaluate data mining of a diverse content corpus if an integration
against an API provides additional metadata for only, say, 10\% of that corpus.

These challenges aside, it is clear that bespoke integrations have a clear place
in this data pipeline being applied in a real enterprise setting. Now that
we have a complete set of theoretical stages for \emph{extraction}, the
remaining improves lie now in the \emph{enrichment} and
\emph{feature generation} stages.

\subsection{Improving Enrichment}

In addition to enriching through dereferencing linked entities, it is proposed
to explore the following options:

\begin{itemize}
\item inferring relationships to hypernyms as defined by Wordnet\cite{miller1995wordnet};
\item inferring facts based using rules derirved from expert domain knowledge;
\item using RDFS and OWL to generate new triples with well-established Ontology rules.
\end{itemize}

In the first approach, we can consider a relationship rule such as:

$$
(IRI, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, <\!\!http\!:\!\!//dbpedia.org/resource/Dog\!\!>)
$$

\noindent and \emph{infer} the fact:

$$
(IRI, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, <\!\!http\!:\!\!//dbpedia.org/resource/Animal\!\!>)
$$

\noindent and produce an enriched graph containing all additional facts
inferred in this way.

When dealing with proper nouns and named entities, inferring facts based on
domain knowledge may be more appropriate. For instance, the rule in n3 syntax:

\begin{lstlistings}
  {
    ?article ex:takesPlaceIn ?city .
    ?city a ex:City .
    ?city ex:capitalOf ?country .
  } -> { ?iri ex:takesPlaceIn ?country }
\end{lstlistings}

\noident might be useful ot help cluster articles that take place in the same
country -- even if the countries are not always explicitly mentioned therein.
Such an inference requires knowledge about cities and countries to
write and domain experts for different types of content might be able to offer
more nuanced rules.

An example for BBC content might be to infer that all articles written under
the \emph{Newsround} brand is suitable for children or that programmes that
have broadcast times during the day are also suitable for children.

The third and final proposed improvement makes use of standard tools to find
\emph{closures} using, e.g. RDFS, ontology rules. With this approach, we
can infer that entities that have a given type or class also have their
superclasses and supertypes. This gives us similar inference to hypernyms,
but with knowledge present in well-established ontologies.

An obvious example might where two content items have been identified as
related to the same concept -- so they would be candidates for clustering
together -- but when RDF data are extracted, it is found that two
different URIs have been used for each:

$$
(IRI_1, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)
(IRI_2, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex2:anotherEntity)
$$

In the RDF graph for IRI_2, say, we might find the source had provided an
\texttt{owl:sameAs} assertion such as:

$$
(ex2:anotherEntity, owl:sameAs, ex1:entity)
$$

This is possible in the case where the second item's data source uses its own
set of identifiers for entities, but has chosen to provide equivalences to
a more standard set of identifiers (e.g. DBpedia). With this information,
our data pipeline can infer:

$$
(IRI_1, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)
(IRI_2, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex2:anotherEntity)
(IRI_2, <\!\!http\!:\!\!//dbpedia.org/property/related\!\!>, ex1:entity)
$$

\noident and the feature generation stage might provide the common features
\texttt{dbprop_related_ex1_entity=true} for both content items.

\subsection{Improving Feature Generation}

The feature generation outlined so far relies solely on boolean values
indicating whether or not a given content item is related by some property to
some entity. This recreates the concept of \emph{tagging} where a given
object is either associated or not associated with a series of \emph{tags}.

One of the advantages of the RDF graph model is that we are not constrained
necessarily to properties and attributes directly applicable to the entity. We
could imagine adding a level of indirection to the query in
Listings~\ref{simple-sparql} to create the query in
Listings~\ref{level2-sparql}.

\begin{lstlisting}[label=lst:level2-sparql,caption={Generates field \texttt{content\_?p1\_?p2\_?v} with value \texttt{true}},language=sparql]
SELECT ?p1 ?p2 ?v
WHERE {
  ?iri ?p1 ?o .
  ?o ?p2 ?v .
}
\end{lstlisting}

With the this query, features of the form \texttt{content\_?p1\_?p2\_?v} can
be generated. An example of this might be where a television programme
content item has information about the actors that appeared therein, e.g.
\texttt{ex:hasActor}, and furthermore we have information about those actors
such as where they were born, e.g. \texttt{ex:bornIn}. With the path created
by following both of these predicates, it is possible to create features for
a television programme such as
\texttt{content\_ex:hasActor\_ex:bornIn\_Edinburgh} and we can potentially
find similarity between programmes where the actors were born in the same
city.

Perhaps a third step in the predicate path followed can give us even more
useful features. The example above could be expanded to
\texttt{content\_ex:hasActor\_ex:bornIn\_ex:cityIn\_Scotland} to allow the
more general ability to cluster programmes with Scottish actors, for instance.

Appropriate experimentation should show whether more value is gained by
adding these additional levels of indirection.

\begin{Maximal Data Pipeline}

In this section, a data pipeline was inductively built up from a base,
identify pipeline with suggestions for potential improvements in different
stages. An application of all the ideas discussed so far might look like
that depicated in Figure\ref{fig:maximal-pipeline}.

\begin{sidewaysfigure}[p]
  \begin{center}
    \begin{dot2tex}[dot,options=-t math,autosize,pgf]
      digraph g {
        rankdir=LR;

@@

        RDF -> Features [label="generate"];
      }
    \end{dot2tex}
  \end{center}
  \caption{\label{fig:maximal-pipeline}Maximal Data Pipeline}
\end{sidewaysfigure}
